---
title: "AR_p272"
format: html
editor: visual
---

### Footing and Agreeing to the Trial Balance  

'Foot' means total the balance of the outstanding accounts receivable at year end.  'Agree' means that this balance should be traced to the trial balance.  If these totals do not match, search for a journal entry that is incorrect.   Investigate reconciling items. If the client has journal entries in the accounts receivable account in the general ledger, the auditors will likely want to review the justification for the larger amounts. This means that these journal entries should be fully documented.

Load the relevant clients files into the R workspace.

```{r  error=F, warning=F, message=F, echo=T, eval=T, message=F, warning=F}

rm(list=ls())
library(readr)


## Set the directory for the new files (modify the path as needed)
default_dir <- "/home/westland/audit_analytics_book/audit_simulated_files/"

if (file.exists(default_dir)){
    setwd(default_dir)
} else {
    dir.create(default_dir)
    setwd(default_dir)
}

# the real_world_ar_ledger file 
# also contains the results of confirmation; 
# the auditor takes a sample of client_ar_ledger records, 
# and auditing obtains information from the real world
# which is simulated here by
# left_join(client_ar_ledger[{sample},], real_world_ar_ledger)

real_world_fyear_end_ar_ledger <-
  read.csv("real_world_fyear_end_ar_ledger.csv", na.strings="0", 
           stringsAsFactors=FALSE) 

fyear_end_ar_ledger <- 
  read.csv("fyear_end_ar_ledger.csv", na.strings="0",
           stringsAsFactors=FALSE)

customer_credit_limits <-  
  read.csv("customer_credit_limits.csv", na.strings="0",
           stringsAsFactors=FALSE)

sales_journal <- 
  read.csv("sales_journal.csv", na.strings="0",
           stringsAsFactors=FALSE)

real_world_credit_sales <- 
  read.csv("real_world_credit_sales.csv", na.strings="0",
           stringsAsFactors=FALSE)

real_world_cash_sales <- 
  read.csv("real_world_cash_sales.csv", na.strings="0",
           stringsAsFactors=FALSE)

```

Foot the client_ar_ledger file and agree the total to the Accounts Receivable entry on the Trial Balance.  Compute the Accounts Receivable balances by customer in preparation for confirmations.

```{r  error=F, warning=F, message=F,  eval=T,  fig.cap="Accounts Receivable Balances by Customer"}

library(tidyverse)
library(lubridate)
library(kableExtra)

# set the fiscal year end date to match the dataset
fyear_end <- paste0(format(Sys.Date(), '%Y'), '-12-31')  


## Set the directory for the new files (modify the path as needed)
default_dir <- "/home/westland/audit_analytics_book/audit_simulated_files/"

if (file.exists(default_dir)){
    setwd(default_dir)
} else {
    dir.create(default_dir)
    setwd(default_dir)
}

## create the unpaid_ar
sales_journal <-  
  read.csv("sales_journal.csv", na.strings="0", stringsAsFactors=FALSE) 

sales_journal[is.na(sales_journal)] <-  0

credit_sales_journal <- 
  sales_journal %>% 
  filter(cash_not_ar == 0)
  
unpaid_ar <- 
  credit_sales_journal[
    credit_sales_journal$collection_date >= fyear_end & 
      credit_sales_journal$shipper_date <= fyear_end,]

foot <- 
  unpaid_ar %>% 
  select(sales_extended) %>% 
  sum()


cat("\n\n A/R balance = ",foot)

# create a list of unpaid AR by customer

unpaid_ar_by_cust <- 
  unpaid_ar %>% 
  select(customer_no, invoice_date, invoice_no, sales_extended) %>% 
  group_by(customer_no) %>% 
  select(sales_extended) %>% 
  summarise(customer_ye_balance=sum(sales_extended)) 

unpaid_ar_by_cust %>% 
  kable(caption="Year End A/R by Customer", 
        "latex", booktabs = T) %>% 
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
  

```
