--- 
title: "Fundamentals of Auditing Financial Reports"
subtitle: "Chapter 1"
author: "J. Christopher Westland"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
  word_document: default
  html_document:
    theme: null
always_allow_html: yes
header-includes:
 \usepackage{float}
---



## Auditing

An audit is an independent examination of the records of an organization to ascertain how far the financial statements as well as non-financial disclosures present a true and fair view of the concern. It also provides assurance that the systems of record keeping are well-controlled and accurate as required by law. Auditing has become such a ubiquitous phenomenon in the corporate and the public sector that academics started identifying an "Audit Society".  

## R packages required for this book

The code in the chapters in this book requires R packages that are specified in the `library("package_name")` commands.  These will be packages such as `tidyverse`, `ggplot2`, `lubridate` and `keras`.  Keras is the API for the Tensorflow machine learning language, and requires a separate `keras` installation with `install_keras`; general notes on Tensorflow installation are provided below.

**Package installation:** There are two steps to using a package.  First it must be _installed_, i.e., copied to a location on your computer where R can access it.  Then it must be _loaded_ into the working memory of _R_.   To install, for example the `tidyverse` package, type *install.packages(“tidyverse”)* and then press the *Enter/Return* key.  To load the previously installed package type *library(tidyverse)*.  After these commands, the `tidyverse` package will now be available for use by your program code.

**Tensorflow installation:** Tensorflow is a machine learning package used in this book; commands to Tensorflow are called using the Keras API.  Prior to using the tensorflow R package you need to install a version of TensorFlow on your system using the R `install_tensorflow()` function, which provides an easy to use wrapper for the various steps required to install TensorFlow.  You can also choose to install TensorFlow manually (as described at https://www.tensorflow.org/install/). 

TensorFlow for R is tested and supported on the following 64-bit systems:

1. Ubuntu 16.04 or later
1. Windows 7 or later
1. macOS 10.12.6 (Sierra) or later (no GPU support)

First, install the tensorflow R package from GitHub or the CRAN respository (search to find the site) then, use the `install_tensorflow()` function to install TensorFlow. Note that on Windows you need a working installation of Anaconda.  `install_tensorflow()` is a wraper around `reticulate::py_install`.  

```{r eval=F}

install.packages("tensorflow")
library(tensorflow)
install_tensorflow()

```

You can confirm that the installation succeeded with:

```{r eval=F}

library(tensorflow)
tf$constant("Hellow Tensorflow")
## tf.Tensor(b'Hellow Tensorflow', shape=(), dtype=string)

```

This will provide you with a default installation of TensorFlow suitable for use with the tensorflow R package.  There is much more to Tensorflow, and interested readers should review the materials at https://tensorflow.rstudio.com/ and at https://www.tensorflow.org/  

